### 2023-08-09 oecd_data/weo_oecd_check_cpi.Rmd
oecd consistency checks

---
title: "WEO against OECD - CPI"
author: "Rike"
date: "2023-08-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.dim = c(11, 8))
library(here)
library(ggplot2)
library(data.table)
library(dplyr)
```

## WEO Truth data against OECD

```{r oecd, warning = FALSE}
oecd <- read.csv(here("oecd_data", "oecd_cpi.csv"))

oecd_q <- oecd |> 
  filter(FREQUENCY == "Q",
         SUBJECT == "TOT") |>
  filter(LOCATION %in% c("CAN", "FRA", "DEU", "ITA", "JPN", "GBR", "USA")) |>
  select(-c(INDICATOR, SUBJECT, FREQUENCY, Flag.Codes)) |>
  rename(country = LOCATION,
         measure = MEASURE,
         yrq = TIME,
         value = Value) |>
  mutate(target_year = substr(yrq, 1, 4) |> as.numeric(),
         quarter = substr(yrq, 7, 7) |> as.numeric()) |>
  filter(target_year >= 1989,
         quarter == 4,
         measure == "AGRWTH") |>
  select(-c(yrq, quarter)) |>
  rename(q4val = value,
         ccode = country) |>
  select(-measure)

oecd_y <- oecd |> 
  filter(FREQUENCY == "A",
         SUBJECT == "TOT") |>
  filter(LOCATION %in% c("CAN", "FRA", "DEU", "ITA", "JPN", "GBR", "USA"),
         MEASURE == "AGRWTH") |>
  select(-c(INDICATOR, SUBJECT, FREQUENCY, Flag.Codes)) |>
  rename(country = LOCATION,
         measure = MEASURE,
         target_year = TIME,
         value = Value) |>
  mutate(target_year = as.numeric(target_year)) |>
  filter(target_year >= 1990) |>
  group_by(country) |>
  arrange(target_year,  .by_group = TRUE) |>
  rename(yearval = value,
         ccode = country) |>
  select(-measure)

weodat <- fread(here("WEOforecasts_tidy.csv"))

weodat_g7 <- weodat |>
  filter(horizon == 0.5,
         !is.na(tv_1),
         g7 == "G7",
         target == "pcpi_pch") |>
  rename(ccode = ISOAlpha_3Code) |>
  select(ccode, country, target_year, tv_1, tv_2) |>
  distinct() |>
  left_join(oecd_y, by = c("ccode", "target_year")) |>
  left_join(oecd_q, by = c("ccode", "target_year")) |>
  setDT() |>
  melt(id.vars = c("ccode", "country", "target_year"), 
       variable.name = "truth_type", 
       value.name = "truth_value") |>
  filter(truth_type != "q4val") |>
  mutate(truth_type = factor(truth_type, 
                             levels = c("tv_1", "tv_2", "yearval", "q4val"),
                             labels = c("WEO_1yr", "WEO_2yr", "OECD_annual", "OECD_q4q4")))

ltdict <- c("dashed", "solid", "solid", "dashed")
coldict <- RColorBrewer::brewer.pal(3, "Set2")[2:3] |> rep(each = 2)
names(ltdict) <- unique(weodat_g7$truth_type)
names(coldict) <- unique(weodat_g7$truth_type)

ggplot(weodat_g7, aes(x = target_year, y = truth_value)) +
  geom_line(aes(linetype = truth_type, color = truth_type)) +
  scale_color_manual(values = coldict) +
  scale_linetype_manual(values = ltdict) +
  facet_wrap(~country, ncol = 2) +
  theme_minimal() %+replace%
  theme(legend.title = element_blank(),
        legend.position = "bottom") +
  ylab("CPI Inflation") +
  xlab("Target Year")
```

### Deviations

Show deviations that are larger than 10% and have absolute deviations of at least 0.3 (to not give "undue attention" to large relative deviations around zero).
```{r pressure, echo=FALSE}
devtab <- weodat_g7 |>
  dcast(ccode + country + target_year ~ truth_type,
        value.var = "truth_value") |>
  select(country, target_year, WEO_2yr, OECD_annual) |>
  mutate(dev = WEO_2yr - OECD_annual,
         dev_pch = 100*(WEO_2yr - OECD_annual)/OECD_annual) |>
    filter(dev_pch > 10 | dev_pch < -10) |>
  filter(dev > 0.3 | dev < -0.3) |>
  mutate(dev = round(dev, 2),
         WEO_2yr = round(WEO_2yr, 2),
         OECD_annual = round(OECD_annual, 2),
         dev_pch = round(dev_pch, 2))
devtab 
```
Some points:

* `r round(100*nrow(devtab)/(nrow(weodat_g7)/4), 2)` % of country-year combinations are affected by 'large' deviations (as defined above) in the two truth data sources
* Deviations clearly affect some countries more than others
  + USA, UK, Canada are not affected at all
  + Germany and the UK are most affected
* Truth data overall seems to be very similar
* annual OECD data more in line with truth values in WEO data than the Q4Q4 OECD data
*