### 2023-09-06 manuscript/manuscript.Rmd
visualize scores and coverage

---
title: "IMF and Benchmark Forecasts"
header-includes:
    - \usepackage{lineno}
    - \linenumbers
    -  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
output: 
  bookdown::pdf_document2:
    extra_dependencies: ["float"]
    keep_tex:  true
    toc: false
linestretch: 2 
bibliography:
  - references.bib
csl: vancouver.csl
---

```{r setup, include=FALSE}
library(data.table)
library(here)
library(ggplot2)
library(patchwork)
library(MetBrewer)
library(purrr)
library(lattice)
library(gt)
library(rlang)
library(tidyverse)
devtools::load_all()
knitr::opts_chunk$set(echo = TRUE)
.d <- `[`
```

# A short note on error handling

In almost all 72 cases, absolute error handling gives lower scores than directional error handling. The only exception is the inflation series for the IMF forecasts and horizon 0, where the expanding window and rolling window method give \textit{slightly} lower scores for the directional methodology. 
We thus decide to focus on the absolute errors in this document.

# Scores, by estimation method, Horizon and forecast source 
```{r, echo = FALSE}
scores <- fread(here("quantile_forecasts", "ci_scores.csv"))

wis_scores <- fread(here("quantile_forecasts", "ci_scores_avgcnt.csv")) |>
  data.table::copy() |>
  .d(, .(model, error_method, method, target, horizon, interval_score, dispersion, underprediction, overprediction)) |>
  setnames("model", "source")

crps_scores <- fread(here("quantile_forecasts", "sample_scores.csv"))

all_scores <- 
  crps_scores[wis_scores, on = c("source", "error_method", "method", "target", "horizon")]|>
  .d(error_method == "absolute") |>
  .d(, error_method := NULL)

gather_scores <- all_scores |>
  data.table::copy() |>
  .d(target == "pcpi_pch") |>
  .d(,target := NULL) |>
  .d(, .(horizon, method, source, score, interval_score)) |>
  setnames("score", "sample_crps") |>
  melt(id.vars = c("horizon", "method", "source"), variable.name = "score_type", value.name = "score") |>
  .d(, combms := paste0(method,"_", score_type)) |>
  .d(, .(horizon, source, combms, score)) |>
  .d(, horizon := paste0("horizon = ", horizon)) |>
  .d(, score := round(score, 3)) |>
  dcast(horizon + combms ~ source, value.var = "score")

gather_scores |>
  data.table::copy() |>
  gt(rowname_col = "combms", groupname_col = "horizon") |>
  cols_label(
      IMF = "IMF",
      ar = "ar",
      bvar = "bvar"
    )

col_names <- c("IMF", "ar", "bvar")

```

## Inflation

```{r, echo = FALSE, warning=FALSE,fig.dim = c(10, 12)}

plots_is <- vector(mode = "list", length = 4)
plots_cr <- vector(mode = "list", length = 4)

for(i in 1:4){

  hor <- (i-1)*0.5

  mydat <- gather_scores |>
    .d(horizon == paste0("horizon = ", hor)) |>
  .d(grep("interval", combms)) |>
  melt(id.vars = c("horizon", "combms"), variable.name = "source", value.name = "score") |>
  .d(, c("method", "scorem", "extra") := tstrsplit(combms, "_", fixed=TRUE))

  plots_is[[i]] <- ggplot(mydat, aes(x = source, y = method)) +
  geom_tile(aes(fill = score)) +
    guides(fill = FALSE) +
    ggtitle(paste0("horizon = ", hor, " - Weighted Interval Score"))
}

for(i in 1:4){

  hor <- (i-1)*0.5

  mydat <- gather_scores |>
    .d(horizon == paste0("horizon = ", hor)) |>
  .d(grep("crps", combms)) |>
  melt(id.vars = c("horizon", "combms"), variable.name = "source", value.name = "score")|>
  .d(, c("method", "scorem", "extra") := tstrsplit(combms, "_", fixed=TRUE))

  plots_cr[[i]] <- ggplot(mydat, aes(x = source, y = method)) +
  geom_tile(aes(fill = score))+
    guides(fill = FALSE)+
    ggtitle(paste0("horizon = ", hor, " - CRPS by sample"))
}

ovr_plot <- 
  (plots_is[[1]] + plots_cr[[1]]) /
  (plots_is[[2]] + plots_cr[[2]]) /
  (plots_is[[3]] + plots_cr[[3]]) /
  (plots_is[[4]] + plots_cr[[4]]) +
  plot_layout(heights = c(1, 1,1,1)) &
    plot_annotation(tag_levels = 'I')  &
    theme(legend.position = 'right', 
          legend.box="vertical", legend.margin=margin())

ovr_plot

```

```{r cvgdat2, include=FALSE, echo = FALSE}

scores <- fread(here("quantile_forecasts", "ci_scores.csv"))
scores_cvgshort <- fread(here("quantile_forecasts", "cvg_pooled.csv"))

cvg_rg <- c(50, 80)

cols <- paste0("coverage_", cvg_rg)

large_cvgdat <- scores |>
  setnames("model", "source") |>
  .d(, c("country", "horizon", "target", "error_method", "method", "source", cols), with = FALSE) |>
  melt(id.vars = c("country", "horizon", "target", "error_method", "method", "source"),
       variable.name = "pilvl",
       value.name = "coverage") |>
  .d(, pilvl := gsub("^.*?coverage_","",pilvl)) |>
  .d(, pilvl := as.numeric(pilvl)/100) |>
  .d(, idcol := paste0(horizon, country, method)) |>
  .d(, country := NULL) |>
  .d(, horizon := NULL)

cvgdat <- scores_cvgshort |>
  setnames("model", "source") |>
  .d(, c( "target", "error_method", "method", "source", cols), with = FALSE) |>
  melt(id.vars = c("target", "error_method", "method", "source"),
       variable.name = "pilvl",
       value.name = "coverage") |>
  .d(, pilvl := gsub("^.*?coverage_","",pilvl)) |>
  .d(, pilvl := as.numeric(pilvl)/100)

cvgdat_cpi <- cvgdat |>
  .d(target == "pcpi_pch")

cvgdat_gdp <- cvgdat |>
  .d(target == "ngdp_rpch")

plot_cpi <- ggplot() +
  geom_line(aes(x = pilvl,
                y = coverage,
                group = idcol),
            data = large_cvgdat |> .d(target == "pcpi_pch"),
            alpha = 0.4,
            color = "gray") +
  geom_point(aes(x = pilvl,
                 y = coverage,
                 color = method),
             data = cvgdat_cpi,
             size = 3) +
  geom_line(aes(x = pilvl,
                y = coverage,
                color = method),
            data = cvgdat_cpi,
            lwd = 1.25) +
  geom_segment(aes(x = 0, y = 0, xend = 1, yend = 1),
               color = "black",
               linetype = "dashed",
               data = cvgdat_cpi) +
  #scale_color_brewer(palette = "Set1") +
  facet_grid(source ~ error_method)+

  ggtitle("Inflation") +

  scale_color_met_d("Hokusai1") +

  ylab("Empirical Coverage Level (Central PI)") +
  xlab("Nominal Coverage (Central PI)") +
  theme_uqimf()

plot_gdp <- ggplot() +
  geom_line(aes(x = pilvl,
                y = coverage,
                group = idcol),
            data = large_cvgdat |> .d(target == "ngdp_rpch"),
            alpha = 0.4,
            color = "gray") +
  geom_point(aes(x = pilvl,
                 y = coverage,
                 color = method),
             data = cvgdat_gdp,
             size = 3) +
  geom_line(aes(x = pilvl,
                y = coverage,
                color = method),
            data = cvgdat_gdp,
            lwd = 1.25) +
  geom_segment(aes(x = 0, y = 0, xend = 1, yend = 1),
               color = "black",
               linetype = "dashed",
               data = cvgdat_gdp) +
  #scale_color_brewer(palette = "Set1") +
  facet_grid(source ~ error_method)+

  ggtitle("Real GDP Growth") +

  scale_color_met_d("Hokusai1") +

  ylab("Empirical Coverage Level (Central PI)") +
  xlab("Nominal Coverage (Central PI)") +
  theme_uqimf()

```

## GDP

```{r hey3, echo = FALSE}
scores <- fread(here("quantile_forecasts", "ci_scores.csv"))

wis_scores <- fread(here("quantile_forecasts", "ci_scores_avgcnt.csv")) |>
  data.table::copy() |>
  .d(, .(model, error_method, method, target, horizon, interval_score, dispersion, underprediction, overprediction)) |>
  setnames("model", "source")

crps_scores <- fread(here("quantile_forecasts", "sample_scores.csv"))

all_scores <- 
  crps_scores[wis_scores, on = c("source", "error_method", "method", "target", "horizon")]|>
  .d(error_method == "absolute") |>
  .d(, error_method := NULL)

gather_scores <- all_scores |>
  data.table::copy() |>
  .d(target == "ngdp_rpch") |>
  .d(,target := NULL) |>
  .d(, .(horizon, method, source, score, interval_score)) |>
  setnames("score", "sample_crps") |>
  melt(id.vars = c("horizon", "method", "source"), variable.name = "score_type", value.name = "score") |>
  .d(, combms := paste0(method,"_", score_type)) |>
  .d(, .(horizon, source, combms, score)) |>
  .d(, horizon := paste0("horizon = ", horizon)) |>
  .d(, score := round(score, 3)) |>
  dcast(horizon + combms ~ source, value.var = "score")

gather_scores |>
  data.table::copy() |>
  gt(rowname_col = "combms", groupname_col = "horizon") |>
  cols_label(
      IMF = "IMF",
      ar = "ar",
      bvar = "bvar"
    )

col_names <- c("IMF", "ar", "bvar")

```

```{r, echo = FALSE, warning=FALSE,fig.dim = c(10, 12)}

plots_is <- vector(mode = "list", length = 4)
plots_cr <- vector(mode = "list", length = 4)

for(i in 1:4){

  hor <- (i-1)*0.5

  mydat <- gather_scores |>
    .d(horizon == paste0("horizon = ", hor)) |>
  .d(grep("interval", combms)) |>
  melt(id.vars = c("horizon", "combms"), variable.name = "source", value.name = "score") |>
  .d(, c("method", "scorem", "extra") := tstrsplit(combms, "_", fixed=TRUE))

  plots_is[[i]] <- ggplot(mydat, aes(x = source, y = method)) +
  geom_tile(aes(fill = score)) +
    guides(fill = FALSE) +
    ggtitle(paste0("horizon = ", hor, " - Weighted Interval Score"))
}

for(i in 1:4){

  hor <- (i-1)*0.5

  mydat <- gather_scores |>
    .d(horizon == paste0("horizon = ", hor)) |>
  .d(grep("crps", combms)) |>
  melt(id.vars = c("horizon", "combms"), variable.name = "source", value.name = "score")|>
  .d(, c("method", "scorem", "extra") := tstrsplit(combms, "_", fixed=TRUE))

  plots_cr[[i]] <- ggplot(mydat, aes(x = source, y = method)) +
  geom_tile(aes(fill = score))+
    guides(fill = FALSE)+
    ggtitle(paste0("horizon = ", hor, " - CRPS by sample"))
}

ovr_plot <- 
  (plots_is[[1]] + plots_cr[[1]]) /
  (plots_is[[2]] + plots_cr[[2]]) /
  (plots_is[[3]] + plots_cr[[3]]) /
  (plots_is[[4]] + plots_cr[[4]]) +
  plot_layout(heights = c(1, 1,1,1)) &
    plot_annotation(tag_levels = 'I')  &
    theme(legend.position = 'right', 
          legend.box="vertical", legend.margin=margin())

ovr_plot

```

```{r cvgdat, include=FALSE, echo = FALSE}

scores <- fread(here("quantile_forecasts", "ci_scores.csv"))
scores_cvgshort <- fread(here("quantile_forecasts", "cvg_pooled.csv"))

cvg_rg <- c(50, 80)

cols <- paste0("coverage_", cvg_rg)

large_cvgdat <- scores |>
  setnames("model", "source") |>
  .d(, c("country", "horizon", "target", "error_method", "method", "source", cols), with = FALSE) |>
  melt(id.vars = c("country", "horizon", "target", "error_method", "method", "source"),
       variable.name = "pilvl",
       value.name = "coverage") |>
  .d(, pilvl := gsub("^.*?coverage_","",pilvl)) |>
  .d(, pilvl := as.numeric(pilvl)/100) |>
  .d(, idcol := paste0(horizon, country, method)) |>
  .d(, country := NULL) |>
  .d(, horizon := NULL)

cvgdat <- scores_cvgshort |>
  setnames("model", "source") |>
  .d(, c( "target", "error_method", "method", "source", cols), with = FALSE) |>
  melt(id.vars = c("target", "error_method", "method", "source"),
       variable.name = "pilvl",
       value.name = "coverage") |>
  .d(, pilvl := gsub("^.*?coverage_","",pilvl)) |>
  .d(, pilvl := as.numeric(pilvl)/100)

cvgdat_cpi <- cvgdat |>
  .d(target == "pcpi_pch")

cvgdat_gdp <- cvgdat |>
  .d(target == "ngdp_rpch")

plot_cpi <- ggplot() +
  geom_line(aes(x = pilvl,
                y = coverage,
                group = idcol),
            data = large_cvgdat |> .d(target == "pcpi_pch"),
            alpha = 0.4,
            color = "gray") +
  geom_point(aes(x = pilvl,
                 y = coverage,
                 color = method),
             data = cvgdat_cpi,
             size = 3) +
  geom_line(aes(x = pilvl,
                y = coverage,
                color = method),
            data = cvgdat_cpi,
            lwd = 1.25) +
  geom_segment(aes(x = 0, y = 0, xend = 1, yend = 1),
               color = "black",
               linetype = "dashed",
               data = cvgdat_cpi) +
  #scale_color_brewer(palette = "Set1") +
  facet_grid(source ~ error_method)+

  ggtitle("Inflation") +

  scale_color_met_d("Hokusai1") +

  ylab("Empirical Coverage Level (Central PI)") +
  xlab("Nominal Coverage (Central PI)") +
  theme_uqimf()

plot_gdp <- ggplot() +
  geom_line(aes(x = pilvl,
                y = coverage,
                group = idcol),
            data = large_cvgdat |> .d(target == "ngdp_rpch"),
            alpha = 0.4,
            color = "gray") +
  geom_point(aes(x = pilvl,
                 y = coverage,
                 color = method),
             data = cvgdat_gdp,
             size = 3) +
  geom_line(aes(x = pilvl,
                y = coverage,
                color = method),
            data = cvgdat_gdp,
            lwd = 1.25) +
  geom_segment(aes(x = 0, y = 0, xend = 1, yend = 1),
               color = "black",
               linetype = "dashed",
               data = cvgdat_gdp) +
  #scale_color_brewer(palette = "Set1") +
  facet_grid(source ~ error_method)+

  ggtitle("Real GDP Growth") +

  scale_color_met_d("Hokusai1") +

  ylab("Empirical Coverage Level (Central PI)") +
  xlab("Nominal Coverage (Central PI)") +
  theme_uqimf()

```

# Coverage, by target, methods and source

```{r cvgplot, fig.dim = c(10, 12), echo = FALSE}

ovr_plot <- 
  (plot_gdp) /
  (plot_cpi) +
  plot_layout(guides = "collect", 
              heights = c(1, 1)) &
    plot_annotation(tag_levels = 'I')  &
    theme(legend.position = 'bottom', 
          legend.box="vertical", legend.margin=margin())

ovr_plot

```

```{r loaddata, include = FALSE, echo = FALSE}

sub_weodat <- 
   data.table::fread(here("WEOforecasts_tidy.csv")) |>
      .d(g7 == 1,) |>
      .d(horizon < 2) |> ######################this shall be extended sometime##############
      .d(, error := prediction - tv_1) |>
      .d(target_year < 2015) |>
      .d(, .(country, prediction, target, error,
             horizon, target_year, forecast_year,
             tv_1)
      )

```

```{r, include = FALSE, echo = FALSE}
head(sub_weodat)
```