### 2023-10-17 README.Rmd
test update Readme

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# uqimf

<!-- badges: start -->
<!-- badges: end -->

The goal of uqimf is to ...

## Contents of the repository
The folders of this repository contain the following:
* `R`: all functions and that the analysis is based on
  - `plot.R`: plotting files 
  - `score.R`: functions to score quantile forecasts (coverage, interval_score, crps by sample)
  - `uqfc.R`: functions to generate quantile forecasts from past errors
  - `uqimf-package.R` some stuff, as this is technically an R package
  - `utils.R` miscellaneous small functions
  - `utils-manuscript.R` miscellaneous small functions for manuscript

```{r cars}
summary(cars)
```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.

You can also embed plots, for example:

```{r pressure, echo = FALSE}
plot(pressure)
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.