### 2023-10-17 README.Rmd
update Readme

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# uqimf

<!-- badges: start -->
<!-- badges: end -->

This repo contains code to generate simple and transparent prediction intervals for the IMF World Economic Outlook forecasts [1]. Central prediction intervals are generated at the 50% and 80% levels for CPI inflation and real GDP growth, based on past forecast errors. See below for documentation on which files the respective folders contain.

[1]: International Monetary Fund. 2023. \textit{World Economic Outlook: A Rocky Recovery.} Washington, DC. April.

## Contents of the repository

The folders of this repository contain the following:

- `R`: All functions that the analysis is based on. Scripts that process data based on these functions are in `scripts`

  - `plot.R`: plotting files

  - `score.R`: functions to score quantile forecasts (coverage, interval_score, crps by sample)

  - `uqfc.R`: functions to generate quantile forecasts from past errors

  - `uqimf-package.R`: some stuff, as this is technically an R package

  - `utils.R`: miscellaneous small functions

  - `utils-manuscript.R`: miscellaneous small functions for manuscript

- `benchmarks`: benchmark forecasts and the scripts that generate them.

  - `point_benchmarks_processed.csv`: point benchmark forecasts (AR and BVAR model) as generated by `scripts/process-benchmarks`, see below

  - `quantile_benchmarks_processed.csv`: quantile benchmark forecasts (BVAR model) as generated by `scripts/process-benchmarks`, see below

- `data`: WEO and merged data

  - `weodat.csv`: tidy IMF WEO data, for G7 countries only
  - `point_forecasts.csv`: all point forecasts (weo and benchmarks), generated by `scripts/process-benchmarks`, see below

- `manuscript`: files to write manuscript. Abandoned now, in favor of `uqimf-manuscript` repository

- `oecd_data`: contains annual and quarterly truth data for inflation and GDP growth, as published by the OECD. Also contains R markdown scripts that check (and verify) compatibility of annual OECD truth values with annual IMF truth values

- `plots`: you get what you bargain for

- `presentation`: some files for the presentation of the project at the MathSEE Symposium

- `quantile-forecasts`: contains quantile forecasts as produced by `scripts/make-forecasts`, see below

- `scores`: Contains scores of forecasts in `quantile-forecasts` and BVAR benchmarks, at various levels of resolutions. Contains its own Readme file to explain what scores the different files contain.

- `scripts`: Contains scripts that process data based on the function in `R`

  - `toy-scripts`: small scripts to explore data

  - `plot-scripts`: scripts to generate plots

  - `download_data.R`: download IMF data

  - `make-forecasts.R`: generate quantile forecasts based (`data/point-forecasts.csv` to `quantile_forecasts/quantile_forecasts.csv`)

  - `process-benchmarks.R` process benchmarks to bring into compatible format with `weodat.csv`. Generates file that contains point forecasts from all sources in long format (`data/point_forecasts.csv`)

  - `process-weo.R`: short script to filter and process WEO data

  - `score-forecasts.R`: score quantile forecasts, those generated here and the benchmark BVAR quantile forecasts.

- `specs`: contains file to specify global settings (window length, etc.)