### 2024-08-29 miscellaneous/exploratory_scripts/length_ci_dirvsabs.Rmd

---
title: "Absolute vs. Directional"
author: "Rike"
date: "2023-11-09"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(data.table)
```

## In the limit: no difference
```{r cars, echo = FALSE}
n <- 10000
dat <- rnorm(n)
dat <- data.table(x = dat)
ggplot(data = dat, aes(x = x)) +
  geom_histogram(color = "steelblue4", fill = "steelblue3") +
  theme_minimal() +
  geom_vline(aes(xintercept = quantile(dat$x, 0.1)), color = "firebrick") + 
  geom_vline(aes(xintercept = quantile(dat$x, 0.9)), color = "firebrick") +
  geom_vline(aes(xintercept = quantile(abs(dat$x), 0.8)), color = "grey20") +
  geom_vline(aes(xintercept = -1*quantile(abs(dat$x), 0.8)), color = "grey20")

```

## Small sample and not symmetric around zero: difference

```{r, echo = FALSE}
library(data.table)
nits <- 100000
n <- 9
set.seed(45)
res15 <- vector(mode = "list", length = nits)
stopp <- TRUE
par(mfrow = c(2,3)) 
k <- 0
resdat <- vector(mode = "list", length = 6)
while(stopp){

  resvec <- c(NA, NA)
  names(resvec) <- c("directional", "absolute")

  dat <- rnorm(n = n)

  resvec[1] <- quantile(dat, 0.9) - quantile(dat, 0.1)
  resvec[2] <- 2*quantile(abs(dat), 0.8)

  if(resvec[2] > resvec[1] + 0.1){
    resdat[[k+1]] <- dat  
    k <- k + 1
  }
  if(k == 6){
    stopp <- FALSE
  }
}

dat <- data.table(x = resdat[[2]])
ggplot(data = dat, aes(x = x)) +
  geom_histogram(bins = 4, color = "steelblue4", fill = "steelblue3") +
  theme_minimal() +
  geom_vline(aes(xintercept = quantile(dat$x, 0.1)), color = "firebrick") + 
  geom_vline(aes(xintercept = quantile(dat$x, 0.9)), color = "firebrick") +
  geom_vline(aes(xintercept = quantile(dat$x, 0.1, type = 8)), color = "grey40") + 
  geom_vline(aes(xintercept = quantile(dat$x, 0.9, type = 8)), color = "grey40") +
  geom_vline(aes(xintercept = quantile(dat$x, 0.1, type = 1)), color = "chartreuse4") + 
  geom_vline(aes(xintercept = quantile(dat$x, 0.9, type = 1)), color = "chartreuse4") +
  geom_vline(aes(xintercept = quantile(dat$x, 0.9)), color = "firebrick") +
  geom_vline(aes(xintercept = quantile(abs(dat$x), 0.8)), color = "red") +
  geom_vline(aes(xintercept = -1*quantile(abs(dat$x), 0.8)), color = "red") +
  geom_vline(aes(xintercept = quantile(abs(dat$x), 0.8, type = 8)), color = "grey10") +
  geom_vline(aes(xintercept = -1*quantile(abs(dat$x), 0.8, type = 8)), color = "grey10") +
  geom_vline(aes(xintercept = quantile(abs(dat$x), 0.8, type = 1)), color = "chartreuse2") +
  geom_vline(aes(xintercept = -1*quantile(abs(dat$x), 0.8, type = 1)), color = "chartreuse2") +
  geom_vline(aes(xintercept = 0), color = "black")

```

## Simulation with 9 samples from normal distribution

```{r, echo = FALSE}
nits <- 10000
n <- 9

res16 <- vector(mode = "list", length = nits)
for(i in 1:nits){

  resvec <- c(NA, NA)
  names(resvec) <- c("directional", "absolute")

  dat <- rnorm(n = n)

  resvec[1] <- quantile(dat, 0.9) - quantile(dat, 0.1)
  resvec[2] <- 2* quantile(abs(dat), 0.8)

  res16[[i]] <- data.table(t(resvec))
}
res16 <- data.table::rbindlist(res16)

res16$diff <- (res16$absolute - res16$directional) > 0

```
```{r, echo = FALSE}
summary(res16$diff)
```
## Simulation with 150 samples from normal distribution
```{r, echo = FALSE}
nits <- 10000
n <- 20000

res16 <- vector(mode = "list", length = nits)
for(i in 1:nits){

  resvec <- c(NA, NA)
  names(resvec) <- c("directional", "absolute")

  dat <- rnorm(n = n)

  resvec[1] <- quantile(dat, 0.9) - quantile(dat, 0.1)
  resvec[2] <- 2* quantile(abs(dat), 0.8)

  res16[[i]] <- data.table(t(resvec))
}
res16 <- data.table::rbindlist(res16)

res16$diff <- (res16$absolute - res16$directional) > 0

```
```{r, echo = FALSE}
summary(res16$diff)
```