### 2024-09-03 README.Rmd
Update Readme

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Simple Macroeconomic Forecast Distributions for the G7 Economies

<!-- badges: start -->
<!-- badges: end -->

This repository contains the documentation, results, and code of a project where we produce simple macroeconomic forecast distributions for the G7 economies, in the form of prediction intervals. It also contains code to continue to produce these forecasts in real-time. See the documentation below for further details.

## Short summary of the project

We generate easy-to-understand forecast intervals for output growth and inflation in the G7 economies, using the point forecasts published in the International Monetary Fund World Economic Outlook (IMF WEO). The forecast intervals are generated at the 50% and 80% levels in a simple and transparent manner, using empirical quantiles of the past history of point forecast errors in the IMF WEO. For more details, please refer to the working paper below.

## Citation

The working paper can be found on arXiv. Please cite this using the following:

> Becker, Krüger, and Schienle. 2024 “Simple Macroeconomic Forecast Distributions for the G7 Economies” arXiv. https://doi.org/10.48550/arXiv.2408.08304.

```
@unpublished{Beckeretal_2024,
  title={Simple Macroeconomic Forecast Distributions for the G7 Economies},
  author={Becker, Friederike and Kr{\"u}ger, Fabian and Schienle, Melanie},
  note={Preprint, arXiv:2408.08304},
  year={2024}
}
```

## Further links

This repository mainly contains files for the retrospective analysis in the working paper, see above. Please visit the following repository to find the forecasts as they are published in real-time, as well as further links to a dashboard, a preregistration protocol, etc.

> https://github.com/KITmetricslab/MacroPI

## Contents of this repository

More details on each folder can be found further below.

Folder | Purpose
---|---
[`benchmarks`](benchmarks/) | Code and forecast data of benchmark forecasts.
[`data`](data/) | Processed WEO and benchmark forecast data. from `data-raw` ready to be used in the paper analysis.
[`miscellanous`](miscellanous/) | Files that are not relevant for the final analysis (exploratory scripts, presentation files, etc). Kept for transparency.
[`oecd_data`](oecd_data/) | OECD truth data used for fitting benchmarks and for robustness checks.
[`quantile_forecasts`](quantile_forecasts/) | The generated forecast intervals, recorded via their upper and lower endpoints. 
[`R`](R/) | R functions used in the analysis and for evaluation.
[`scores`](scores/) | Scores of forecasts.
[`scripts`](scripts/) | Executive scripts to produce all results.
[`specs`](specs/) | Specifications and global options for the analysis.

Forecast intervals at the 50% and 80% are recorded via their upper and lower endpoints, that is, the 0.25 and 0.75 predictive quantiles for the 50% interval, and the 0.1 and 0.9 predictive quantiles for the 90% interval.

The folders of this repository contain the following:

- `R`: All functions that the analysis is based on. Scripts that process data based on these functions are in `scripts`

  - `plot.R`: plotting files

  - `score.R`: functions to score quantile forecasts (coverage, interval_score, crps by sample)

  - `uqfc.R`: functions to generate quantile forecasts from past errors

  - `uqimf-package.R`: some stuff, as this is technically an R package

  - `utils.R`: miscellaneous small functions

  - `utils-manuscript.R`: miscellaneous small functions for manuscript

- `benchmarks`: benchmark forecasts and the scripts that generate them.

  - `point_benchmarks_processed.csv`: point benchmark forecasts (AR and BVAR model) as generated by `scripts/process-benchmarks`, see below

  - `quantile_benchmarks_processed.csv`: quantile benchmark forecasts (BVAR model) as generated by `scripts/process-benchmarks`, see below

- `data`: WEO and merged data

  - `weodat.csv`: tidy IMF WEO data, for G7 countries only
  - `point_forecasts.csv`: all point forecasts (weo and benchmarks), generated by `scripts/process-benchmarks`, see below

- `manuscript`: files to write manuscript. Abandoned now, in favor of `uqimf-manuscript` repository

- `oecd_data`: contains annual and quarterly truth data for inflation and GDP growth, as published by the OECD. Also contains R markdown scripts that check (and verify) compatibility of annual OECD truth values with annual IMF truth values

- `plots`: you get what you bargain for

- `presentation`: some files for the presentation of the project at the MathSEE Symposium

- `quantile-forecasts`: contains quantile forecasts as produced by `scripts/make-forecasts`, see below

- `scores`: Contains scores of forecasts in `quantile-forecasts` and BVAR benchmarks, at various levels of resolutions. Contains its own Readme file to explain what scores the different files contain.

- `scripts`: Contains scripts that process data based on the function in `R`

  - `toy-scripts`: small scripts to explore data

  - `plot-scripts`: scripts to generate plots

  - `download_data.R`: download IMF data

  - `make-forecasts.R`: generate quantile forecasts based (`data/point-forecasts.csv` to `quantile_forecasts/quantile_forecasts.csv`)

  - `process-benchmarks.R` process benchmarks to bring into compatible format with `weodat.csv`. Generates file that contains point forecasts from all sources in long format (`data/point_forecasts.csv`)

  - `process-weo.R`: short script to filter and process WEO data

  - `score-forecasts.R`: score quantile forecasts, those generated here and the benchmark BVAR quantile forecasts.

- `specs`: contains file to specify global settings (window length, etc.)

## References

International Monetary Fund. 2023. *World Economic Outlook: A Rocky Recovery.* Washington, DC. April.