### 2025-04-06 manuscript_plots/revision/extcis_current.tex
automate more elements of table, fix order of footnotes

\begin{table}[!h]
\centering
\caption{Interval Scores with Minimums Highlighted}
\centering
\begin{tabular}[t]{llllllllllllllll}
\toprule
\multicolumn{2}{c}{ } & \multicolumn{7}{c}{{Inflation\hspace*{15mm}}} & \multicolumn{7}{c}{{GDP Growth}} \\
\cmidrule(l{3pt}r{3pt}){3-9} \cmidrule(l{3pt}r{3pt}){10-16}
 &  & $\text{CRPS}^{1)}$-c & $\text{IS}_{W}^{2)}$-c & $\text{IS}_{W,b}^{2)}$-c & $\text{IS}_{U}^{2)}$-c & $\text{IS}_{50}^{2)}$-c & $\text{IS}_{80}^{2)}$-c & $\text{IS}_{\text{CISS}}^{2)}$-c & $\text{CRPS}^{1)}$-g & $\text{IS}_{W}^{2)}$-g & $\text{IS}_{W,b}^{2)}$-g & $\text{IS}_{U}^{2)}$-g & $\text{IS}_{50}^{2)}$-g & $\text{IS}_{80}^{2)}$-g & $\text{IS}_{\text{CISS}}^{2)}$-g\\
\midrule
\cellcolor{gray!35}{} & \cellcolor{gray!35}{IMF} & \cellcolor{gray!35}{\textbf{0.103}} & \cellcolor{gray!35}{\textbf{0.106}} & \cellcolor{gray!35}{\textbf{0.093}} & \cellcolor{gray!35}{\textbf{0.522}} & \cellcolor{gray!35}{\textbf{0.46}} & \cellcolor{gray!35}{\textbf{0.741}} & \cellcolor{gray!35}{\textbf{0.089}} & \cellcolor{gray!35}{0.292} & \cellcolor{gray!35}{0.298} & \cellcolor{gray!35}{0.269} & \cellcolor{gray!35}{1.561} & \cellcolor{gray!35}{1.279} & \cellcolor{gray!35}{2.298} & \cellcolor{gray!35}{0.255}\\
\parbox[t]{2mm}{\multirow{13}{*}{\rotatebox[origin=c]{90}{\hspace{5mm}Fall, Current}}}
 & AR(1) & 0.195 & 0.198 & 0.177 & 1.034 & 0.839 & 1.484 & 0.214 & 0.337 & 0.344 & 0.298 & 1.697 & 1.485 & 2.355 & 0.302\\
 & AR(p) & 0.183 & 0.185 & 0.166 & 0.978 & 0.783 & 1.449 & 0.203 & 0.333 & 0.34 & 0.296 & 1.688 & 1.466 & 2.39 & 0.3\\
 & BVAR$^{3)}$ & 0.2 & 0.203 & 0.183 & 1.053 & 0.87 & 1.494 & 0.219 & 0.326 & 0.334 & 0.285 & 1.664 & 1.449 & 2.356 & 0.288\\
 & BVAR-Mix$^{3)}$ & 0.198 & 0.201 & 0.182 & 1.042 & 0.86 & 1.47 & 0.218 & 0.323 & 0.33 & 0.284 & 1.642 & 1.417 & 2.311 & 0.287\\
 & Direct$^{4)}$: AR(1) & 0.195 & 0.196 & 0.182 & 1.048 & 0.843 & 1.532 & 0.217 & 0.332 & 0.334 & 0.303 & 1.713 & 1.462 & 2.412 & 0.303\\
 & Direct$^{4)}$: AR(p) & 0.181 & 0.183 & 0.169 & 0.981 & 0.785 & 1.424 & 0.203 & 0.326 & 0.328 & 0.298 & 1.684 & 1.437 & 2.363 & 0.3\\
 & Direct$^{4)}$: AR-annual & 1.03 & 0.988 & 0.899 & 5.179 & 4.252 & 7.355 & 0.987 & 1.711 & 1.613 & 1.56 & 8.995 & 7.043 & 13.599 & 1.678\\
 & Direct$^{4)}$: ARX-annual & 0.133 & 0.134 & 0.121 & 0.653 & 0.616 & 0.879 & 0.121 & 0.33 & 0.324 & 0.297 & 1.67 & 1.435 & 2.359 & 0.301\\
 & Direct$^{4)}$: BVAR$^{3)}$ & 0.2 & 0.201 & 0.184 & 1.069 & 0.859 & 1.537 & 0.221 & 0.328 & 0.33 & 0.303 & 1.735 & 1.425 & 2.491 & 0.305\\
 & Direct$^{4)}$: BVAR-CISS$^{3)}$ & NA & NA & NA & NA & NA & NA & 0.244 & NA & NA & NA & NA & NA & NA & 0.344\\
 & Direct$^{4)}$: BVAR-Mix$^{3)}$ & 0.2 & 0.201 & 0.185 & 1.07 & 0.864 & 1.539 & 0.224 & 0.335 & 0.336 & 0.311 & 1.798 & 1.446 & 2.613 & 0.311\\
\cellcolor{gray!15}{} & \cellcolor{gray!15}{Simple Ensemble$^{5)}$} & \cellcolor{gray!15}{0.145} & \cellcolor{gray!15}{0.147} & \cellcolor{gray!15}{0.132} & \cellcolor{gray!15}{0.752} & \cellcolor{gray!15}{0.636} & \cellcolor{gray!15}{1.063} & \cellcolor{gray!15}{0.156} & \cellcolor{gray!15}{\textbf{0.23}} & \cellcolor{gray!15}{\textbf{0.237}} & \cellcolor{gray!15}{\textbf{0.208}} & \cellcolor{gray!15}{\textbf{1.194}} & \cellcolor{gray!15}{\textbf{1.045}} & \cellcolor{gray!15}{\textbf{1.756}} & \cellcolor{gray!15}{\textbf{0.201}}\\
\addlinespace\addlinespace
\cellcolor{gray!35}{} & \cellcolor{gray!35}{IMF} & \cellcolor{gray!35}{0.419} & \cellcolor{gray!35}{0.425} & \cellcolor{gray!35}{0.387} & \cellcolor{gray!35}{2.25} & \cellcolor{gray!35}{1.807} & \cellcolor{gray!35}{3.257} & \cellcolor{gray!35}{0.425} & \cellcolor{gray!35}{\textbf{0.506}} & \cellcolor{gray!35}{\textbf{0.518}} & \cellcolor{gray!35}{\textbf{0.466}} & \cellcolor{gray!35}{2.64} & \cellcolor{gray!35}{\textbf{2.254}} & \cellcolor{gray!35}{3.958} & \cellcolor{gray!35}{0.496}\\
\parbox[t]{2mm}{\multirow{13}{*}{\rotatebox[origin=c]{90}{\hspace{5mm}Spring, Current}}}
 & AR(1) & 0.454 & 0.46 & 0.405 & 2.319 & 1.958 & 3.243 & 0.441 & 0.816 & 0.829 & 0.742 & 4.292 & 3.487 & 6.189 & 0.699\\
 & AR(p) & 0.464 & 0.47 & 0.438 & 2.401 & 2.024 & 3.395 & 0.473 & 0.739 & 0.751 & 0.676 & 3.91 & 3.186 & 5.617 & 0.649\\
 & BVAR$^{3)}$ & 0.455 & 0.461 & 0.41 & 2.334 & 1.965 & 3.325 & 0.45 & 0.716 & 0.728 & 0.651 & 3.763 & 3.131 & 5.369 & 0.606\\
 & BVAR-Mix$^{3)}$ & 0.458 & 0.465 & 0.412 & 2.365 & 1.977 & 3.342 & 0.453 & 0.697 & 0.709 & 0.637 & 3.66 & 3.06 & 5.175 & 0.594\\
 & Direct$^{4)}$: AR(1) & 0.453 & 0.458 & 0.419 & 2.341 & 2.015 & 3.338 & 0.444 & 0.796 & 0.802 & 0.739 & 4.216 & 3.458 & 6.129 & 0.696\\
 & Direct$^{4)}$: AR(p) & 0.462 & 0.466 & 0.43 & 2.386 & 2.082 & 3.394 & 0.469 & 0.72 & 0.726 & 0.677 & 3.83 & 3.174 & 5.596 & 0.654\\
 & Direct$^{4)}$: AR-annual & 1.03 & 0.988 & 0.899 & 5.178 & 4.25 & 7.352 & 0.987 & 1.714 & 1.617 & 1.563 & 9.008 & 7.057 & 13.625 & 1.681\\
 & Direct$^{4)}$: ARX-annual & 0.474 & 0.462 & 0.424 & 2.402 & 2.024 & 3.413 & 0.476 & 0.529 & \textbf{0.518} & 0.468 & \textbf{2.615} & 2.28 & \textbf{3.667} & 0.511\\
 & Direct$^{4)}$: BVAR$^{3)}$ & 0.456 & 0.462 & 0.426 & 2.358 & 2.064 & 3.352 & 0.458 & 0.69 & 0.695 & 0.637 & 3.625 & 3.026 & 5.164 & 0.59\\
 & Direct$^{4)}$: BVAR-CISS$^{3)}$ & NA & NA & NA & NA & NA & NA & 0.502 & NA & NA & NA & NA & NA & NA & 0.671\\
 & Direct$^{4)}$: BVAR-Mix$^{3)}$ & 0.429 & 0.433 & 0.384 & 2.169 & 1.861 & 3.024 & 0.411 & 0.667 & 0.671 & 0.611 & 3.502 & 2.905 & 4.961 & 0.569\\
\cellcolor{gray!15}{} & \cellcolor{gray!15}{Simple Ensemble$^{5)}$} & \cellcolor{gray!15}{\textbf{0.389}} & \cellcolor{gray!15}{\textbf{0.395}} & \cellcolor{gray!15}{\textbf{0.351}} & \cellcolor{gray!15}{\textbf{1.998}} & \cellcolor{gray!15}{\textbf{1.69}} & \cellcolor{gray!15}{\textbf{2.823}} & \cellcolor{gray!15}{\textbf{0.384}} & \cellcolor{gray!15}{0.559} & \cellcolor{gray!15}{0.571} & \cellcolor{gray!15}{0.508} & \cellcolor{gray!15}{2.893} & \cellcolor{gray!15}{2.474} & \cellcolor{gray!15}{4.154} & \cellcolor{gray!15}{\textbf{0.489}}\\
\bottomrule
\end{tabular}
\end{table}