### 2025-04-06 manuscript_plots/revision/extcis_next.tex

\begin{table}[!h]
\centering
\caption{Interval Scores with Minimums Highlighted}
\centering
\begin{tabular}[t]{llllllllllllllll}
\toprule
\multicolumn{2}{c}{ } & \multicolumn{7}{c}{{Inflation\hspace*{15mm}}} & \multicolumn{7}{c}{{GDP Growth}} \\
\cmidrule(l{3pt}r{3pt}){3-9} \cmidrule(l{3pt}r{3pt}){10-16}
 &  & $\text{CRPS}^{1)}$-c & $\text{IS}_{W}^{2)}$-c & $\text{IS}_{W,b}^{2)}$-c & $\text{IS}_{U}^{2)}$-c & $\text{IS}_{50}^{2)}$-c & $\text{IS}_{80}^{2)}$-c & $\text{IS}_{\text{CISS}}^{2)}$-c & $\text{CRPS}^{1)}$-g & $\text{IS}_{W}^{2)}$-g & $\text{IS}_{W,b}^{2)}$-g & $\text{IS}_{U}^{2)}$-g & $\text{IS}_{50}^{2)}$-g & $\text{IS}_{80}^{2)}$-g & $\text{IS}_{\text{CISS}}^{2)}$-g\\
\midrule
\cellcolor{gray!35}{} & \cellcolor{gray!35}{IMF} & \cellcolor{gray!35}{0.914} & \cellcolor{gray!35}{0.925} & \cellcolor{gray!35}{0.825} & \cellcolor{gray!35}{4.863} & \cellcolor{gray!35}{3.869} & \cellcolor{gray!35}{6.935} & \cellcolor{gray!35}{0.899} & \cellcolor{gray!35}{\textbf{1.173}} & \cellcolor{gray!35}{\textbf{1.189}} & \cellcolor{gray!35}{\textbf{1.128}} & \cellcolor{gray!35}{\textbf{6.475}} & \cellcolor{gray!35}{\textbf{5.105}} & \cellcolor{gray!35}{\textbf{9.732}} & \cellcolor{gray!35}{\textbf{1.188}}\\
\parbox[t]{2mm}{\multirow{13}{*}{\rotatebox[origin=c]{90}{\hspace{5mm}Fall, Next}}}
 & AR(1) & 1.056 & 1.071 & 0.904 & 5.08 & 4.581 & 6.653 & 0.977 & 1.494 & 1.514 & 1.367 & 7.944 & 6.368 & 11.43 & 1.368\\
 & AR(p) & 0.955 & 0.968 & 0.851 & 4.903 & 4.147 & 6.713 & 0.9 & 1.427 & 1.446 & 1.32 & 7.689 & 6.109 & 11.154 & 1.328\\
 & BVAR$^{3)}$ & 1.006 & 1.02 & 0.857 & 4.883 & 4.373 & 6.403 & 0.954 & 1.368 & 1.387 & 1.289 & 7.437 & 5.841 & 11.009 & 1.313\\
 & BVAR-Mix$^{3)}$ & 0.876 & 0.888 & \textbf{0.758} & 4.332 & \textbf{3.776} & 5.85 & 0.854 & 1.334 & 1.352 & 1.26 & 7.269 & 5.699 & 10.737 & 1.287\\
 & Direct$^{4)}$: AR(1) & 1.037 & 1.049 & 0.91 & 5.052 & 4.52 & 6.898 & 0.973 & 1.487 & 1.497 & 1.381 & 7.978 & 6.388 & 11.644 & 1.383\\
 & Direct$^{4)}$: AR(p) & 0.93 & 0.938 & 0.844 & 4.689 & 4.136 & 6.546 & 0.88 & 1.417 & 1.427 & 1.34 & 7.701 & 6.16 & 11.392 & 1.349\\
 & Direct$^{4)}$: AR-annual & 1.374 & 1.301 & 1.194 & 6.913 & 5.582 & 9.935 & 1.354 & 1.545 & 1.476 & 1.424 & 8.127 & 6.49 & 12.253 & 1.524\\
 & Direct$^{4)}$: ARX-annual & 1.129 & 1.06 & 0.976 & 5.613 & 4.588 & 8.054 & 1.093 & 1.569 & 1.49 & 1.381 & 7.969 & 6.417 & 11.578 & 1.523\\
 & Direct$^{4)}$: BVAR$^{3)}$ & 0.995 & 1.007 & 0.887 & 4.903 & 4.389 & 6.76 & 0.952 & 1.361 & 1.37 & 1.29 & 7.458 & 5.904 & 11.035 & 1.31\\
 & Direct$^{4)}$: BVAR-CISS$^{3)}$ & NA & NA & NA & NA & NA & NA & 1.05 & NA & NA & NA & NA & NA & NA & 1.871\\
 & Direct$^{4)}$: BVAR-Mix$^{3)}$ & \textbf{0.857} & \textbf{0.866} & 0.771 & \textbf{4.215} & 3.86 & \textbf{5.766} & \textbf{0.851} & 1.338 & 1.346 & 1.27 & 7.428 & 5.735 & 11.058 & 1.286\\
\cellcolor{gray!15}{} & \cellcolor{gray!15}{Simple Ensemble$^{5)}$} & \cellcolor{gray!15}{0.93} & \cellcolor{gray!15}{0.943} & \cellcolor{gray!15}{0.817} & \cellcolor{gray!15}{4.644} & \cellcolor{gray!15}{4.055} & \cellcolor{gray!15}{6.27} & \cellcolor{gray!15}{0.896} & \cellcolor{gray!15}{1.286} & \cellcolor{gray!15}{1.304} & \cellcolor{gray!15}{1.213} & \cellcolor{gray!15}{7.011} & \cellcolor{gray!15}{5.509} & \cellcolor{gray!15}{10.394} & \cellcolor{gray!15}{1.242}\\
\addlinespace\addlinespace
\cellcolor{gray!35}{} & \cellcolor{gray!35}{IMF} & \cellcolor{gray!35}{1.108} & \cellcolor{gray!35}{1.119} & \cellcolor{gray!35}{1.013} & \cellcolor{gray!35}{5.895} & \cellcolor{gray!35}{\textbf{4.741}} & \cellcolor{gray!35}{8.397} & \cellcolor{gray!35}{1.123} & \cellcolor{gray!35}{\textbf{1.25}} & \cellcolor{gray!35}{\textbf{1.271}} & \cellcolor{gray!35}{\textbf{1.204}} & \cellcolor{gray!35}{\textbf{6.878}} & \cellcolor{gray!35}{\textbf{5.481}} & \cellcolor{gray!35}{\textbf{10.366}} & \cellcolor{gray!35}{\textbf{1.267}}\\
\parbox[t]{2mm}{\multirow{13}{*}{\rotatebox[origin=c]{90}{\hspace{5mm}Spring, Next}}}
 & AR(1) & 1.377 & 1.396 & 1.142 & 6.349 & 5.952 & 7.923 & 1.195 & 1.769 & 1.791 & 1.587 & 9.22 & 7.521 & 13.037 & 1.6\\
 & AR(p) & 1.113 & 1.126 & 0.975 & 5.683 & 4.767 & 8.025 & 1.057 & 1.757 & 1.781 & 1.608 & 9.347 & 7.532 & 13.432 & 1.667\\
 & BVAR$^{3)}$ & 1.346 & 1.363 & 1.122 & 6.304 & 5.764 & 7.996 & 1.215 & 1.686 & 1.708 & 1.571 & 9.124 & 7.245 & 13.281 & 1.635\\
 & BVAR-Mix$^{3)}$ & 1.126 & 1.141 & 0.962 & 5.408 & 4.86 & 7.137 & 1.06 & 1.635 & 1.656 & 1.535 & 8.874 & 7.051 & 12.98 & 1.607\\
 & Direct$^{4)}$: AR(1) & 1.35 & 1.365 & 1.131 & 6.367 & 5.788 & 8.156 & 1.185 & 1.763 & 1.775 & 1.601 & 9.183 & 7.581 & 13.06 & 1.609\\
 & Direct$^{4)}$: AR(p) & \textbf{1.095} & \textbf{1.106} & 1.002 & 5.63 & 4.804 & 8.023 & 1.061 & 1.719 & 1.731 & 1.579 & 9.111 & 7.365 & 13.16 & 1.646\\
 & Direct$^{4)}$: AR-annual & 1.373 & 1.301 & 1.194 & 6.912 & 5.582 & 9.933 & 1.354 & 1.546 & 1.477 & 1.424 & 8.128 & 6.493 & 12.253 & 1.525\\
 & Direct$^{4)}$: ARX-annual & 1.323 & 1.245 & 1.155 & 6.68 & 5.362 & 9.694 & 1.319 & 1.496 & 1.404 & 1.33 & 7.611 & 6.114 & 11.307 & 1.418\\
 & Direct$^{4)}$: BVAR$^{3)}$ & 1.336 & 1.352 & 1.133 & 6.39 & 5.697 & 8.41 & 1.194 & 1.651 & 1.661 & 1.533 & 8.848 & 7.14 & 12.818 & 1.601\\
 & Direct$^{4)}$: BVAR-CISS$^{3)}$ & NA & NA & NA & NA & NA & NA & 1.103 & NA & NA & NA & NA & NA & NA & 1.876\\
 & Direct$^{4)}$: BVAR-Mix$^{3)}$ & \textbf{1.095} & 1.107 & \textbf{0.942} & \textbf{5.254} & 4.804 & \textbf{6.82} & \textbf{1.02} & 1.611 & 1.62 & 1.501 & 8.729 & 6.942 & 12.659 & 1.575\\
\cellcolor{gray!15}{} & \cellcolor{gray!15}{Simple Ensemble$^{5)}$} & \cellcolor{gray!15}{1.18} & \cellcolor{gray!15}{1.196} & \cellcolor{gray!15}{0.993} & \cellcolor{gray!15}{5.617} & \cellcolor{gray!15}{5.072} & \cellcolor{gray!15}{7.168} & \cellcolor{gray!15}{1.069} & \cellcolor{gray!15}{1.509} & \cellcolor{gray!15}{1.53} & \cellcolor{gray!15}{1.403} & \cellcolor{gray!15}{8.122} & \cellcolor{gray!15}{6.436} & \cellcolor{gray!15}{11.922} & \cellcolor{gray!15}{1.455}\\
\bottomrule
\end{tabular}
\end{table}